Cooperative Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-13 manuscript/paper.tex
t2g gen case. figs explained and discussed.

$E_D, E_{\pm}$ relative to their average
and the weight of $xy$ state $W_{xy}$ in $E_{\pm}$,
as a function of $\alpha$ at $q_A/q_o=1/2,1,2$ (and $\Delta=15q_o/4a$)
are shown in Fig.~\ref{fig:t2g-beta0}.
The average shifts slightly with $\alpha,q_A$ by 
$E_0^{t_{2g}}= \left(E_{xy} + 2 E_{yz}\right)/3-(2/5)f_{q_A}\Delta_{O}$,
again reflecting 
a change in the gap $\Delta$,
and is also included in Fig.~\ref{fig:t2g-beta0}.
Apart from the mixing between the three states $xy,yz,zx$ in $\ket{\pm}$ at finite $\alpha$, 
Fig.~\ref{fig:t2g-beta0} shows a similar trend as in Fig.~\ref{fig:t2g-alpha0}
but with an interesting twist.
%Fig.~\ref{fig:evals-alpha} reveals another interesting point.
In contrast to $\alpha=0$ case discussed in previous section,
$W_{xy}$ is the largest in the highest (lowest) energy state at small (large) $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-beta0}(a,c).
$\alpha$ and $\beta$ would thus tend to produce different ordering
(details in the next section).
The splittings are also abut the same magnitude, between $4$-$6\%$ at $\alpha=15^\circ$
at $q_A/q_o=1/2,2$ [Fig.~\ref{fig:t2g-beta0}(a,c)] and about $2\%$ at $q_A/q_o=1$ [Fig.~\ref{fig:t2g-beta0}(b)].
The mixing is sizeable but still far from perfect ($50/50\%$) even at the largest $\alpha$ shown ($W_{xy}=0.25$-$0.35$ at $20^\circ$).
% At unrealistically large $\alpha\sim30-40^\circ$, 
Overall, at finite $\alpha$,
the lowest energy state is either purely $xy/yz$,
or contains at least some fraction of these,
but it is never pure $xy$.

\subsubsection{Rotation and tilting: $\alpha,\beta\neq 0$}

The t2g Hamiltonian in the general case cannot be solve analytically
 so there is no longer any advantage of ignoring the coupling between 
 the eg and t2g manifolds.
We can thus keep these couplings and numerically solve 
the full Hamiltonian instead.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{default}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/wts-energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{default}
\end{center}
\end{figure}

>>>>

$E_D, E_{\pm}$ relative to their average
and the weight of $xy$ state $W_{xy}$ in $E_{\pm}$,
as a function of $\alpha$ at $q_A/q_o=1/2,1,2$ (and $\Delta=15q_o/4a$)
are shown in Fig.~\ref{fig:t2g-beta0}.
The average shifts slightly with $\alpha,q_A$ by 
$E_0^{t_{2g}}= \left(E_{xy} + 2 E_{yz}\right)/3-(2/5)f_{q_A}\Delta_{O}$,
again reflecting 
a change in the gap $\Delta$,
and is also included in Fig.~\ref{fig:t2g-beta0}.
Apart from the mixing between the three states $xy,yz,zx$ in $\ket{\pm}$ at finite $\alpha$, 
Fig.~\ref{fig:t2g-beta0} shows a similar trend as in Fig.~\ref{fig:t2g-alpha0}
but with an interesting twist.
%Fig.~\ref{fig:evals-alpha} reveals another interesting point.
In contrast to $\alpha=0$ case discussed in the previous section,
$W_{xy}$ is the largest in the highest (lowest) energy state at small (large) $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-beta0}(a,c).
$\alpha$ and $\beta$ would thus tend to produce different ordering
(details in the next section).
The splittings are also abut the same magnitude, between $4$-$6\%$ at $\alpha=15^\circ$
at $q_A/q_o=1/2,2$ [Fig.~\ref{fig:t2g-beta0}(a,c)] and about $2\%$ at $q_A/q_o=1$ [Fig.~\ref{fig:t2g-beta0}(b)].
The mixing is sizeable but still far from perfect ($50/50\%$) even at the largest $\alpha$ shown ($W_{xy}=0.25$-$0.35$ at $20^\circ$).
% At unrealistically large $\alpha\sim30-40^\circ$, 
Overall, at finite $\alpha$,
the lowest energy state is either purely $xy/yz$,
or contains at least some fraction of these,
but it is never pure $xy$.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{fig:t2g-ge-Es}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/wts-energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{fig:t2g-ge-wts}
\end{center}
\end{figure}

\subsubsection{Rotation and tilting: $\alpha,\beta\neq 0$}
\label{sec:t2g-gen}

The $t_{2g}$ Hamiltonian in the general case $\alpha,\beta\neq 0$ cannot be solve analytically
 so there is no longer any advantage of ignoring the coupling between 
 the $e_{g}$ and $t_{2g}$ manifolds.
Even though their effect should be negligible due to a large gap $\Delta$ between the two manifolds,
here we keep these couplings 
and numerically solve
the full Hamiltonian.
Figure~\ref{fig:t2g-ge-Es}
shows the energies at $\alpha=10^\circ$
as a function of $\beta$
at $q_A/q_o=1,2$ and $\Delta_o=4q_o/15a$ again.
All states in the $t_{2g}$ mix up to make the eigenstates $\ket{E_i}, i=1,2,3$
at energies $E_i, i=1,2,3$.
\az{E0, can we show E0 only in the first fig and not all of them?}
At $\beta=0$, these results correspond to $\alpha=10^\circ$ 
in Figs.~\ref{fig:t2g-beta0}(a,c),
so the lowest state at $\beta=0$
here is $\ket{D}$ in Figs.~\ref{fig:t2g-ge-Es}(a)
while it is the highest in Figs.~\ref{fig:t2g-ge-Es}(b).
We see two anticrossings or level repulsion in either of Figs.~\ref{fig:t2g-ge-Es}(a,b)
in different orders.
In Figs.~\ref{fig:t2g-ge-Es}(a), 
the anticorssing between the highest two states occurs at smaller $\beta$,
while the situation is the opposite in Figs.~\ref{fig:t2g-ge-Es}(b) where 
it occurs between the lowest pair at smaller $\beta$.
From the results in the previous two sections,
we know that not only A and B but also $\alpha$ and $\beta$ tend to establish different orders.
To get more insight, the weights of these three $t_{2g}$ states in the three eigenstates 
% in Fig.~\ref{fig:t2g-ge-Es}
are shown in Fig.~\ref{fig:t2g-ge-wts}.
Let's first focus on the left column ($q_A/q_o=1/2$). 
Fig.~\ref{fig:t2g-ge-wts}(a)
shows that, while $\ket{E_1}\to \ket{D}$ as $\beta\to 0$,
$\ket{E_1}$ practically stays in $yz,xz$ subspace
for $\beta<7^\circ$ and it quickly becomes $xy$
as $\beta\alpha=10^\circ$.
Fig.~\ref{fig:t2g-ge-wts}(b,c)
show a complimentary behaviour 
between $xy$ and $yz,xz$,
where $\ket{E_3}$
transforms from almost pure $xy$ at $\beta=0$
to almost complete $yz,xz$ at $\beta\gtrsim\alpha=10^\circ$.
The right column
in Fig.~\ref{fig:t2g-ge-wts} ($q_A/q_o=2$),
shows the opposite behaviour 
about the ``resonance'' $\beta\sim\alpha$.
Fig.~\ref{fig:t2g-ge-wts}(b,e)
show an interesting aspect,
the $W_{xy}$ peaking in the middle state $\ket{E_2}$
around before it transfers between $\ket{E_1}$ and $\ket{E_2}$
around $\beta\sim\alpha$,
i.e., at $\beta$
between the two anticrossings in 
Fig.~\ref{fig:t2g-ge-Es},
which makes sense if we imagine the continuation of various energy surfaces/curves
in 
the spectra in Fig.~\ref{fig:t2g-ge-Es}
ignoring their anticorssings:
$xy$ starts from the top and goes down in 
Fig.~\ref{fig:t2g-ge-Es}(a),
while it starts from the bottom and goes up
in Fig.~\ref{fig:t2g-ge-Es}(b).
